---
title: "BOUN Spring 2019 ETM 58D Homework 2"
author: "Adil Arca Seney & Begüm Urgenç"
date: "April 29, 2019"
output: html_document
---
#Task1

```{r}

library(data.table)
library(plotly)
library(ggplot2)
library(ggfortify)
library(scatterplot3d)

matches=readRDS("C:\\Users\\a-adsene\\Desktop\\ETM58D\\df9b1196-e3cf-4cc7-9159-f236fe738215_matches.rds")

odds=readRDS('C:\\Users\\a-adsene\\Desktop\\ETM58D\\df9b1196-e3cf-4cc7-9159-f236fe738215_odd_details.rds')

matches$type=NULL
matches$leagueId=NULL
matches$home=NULL
matches$away=NULL
matches$date=NULL

matches[,c('score_home','score_away') := tstrsplit(score,':')]
matches[,score_home:=as.numeric(score_home)]
matches[,score_away:=as.numeric(score_away)]
matches[,c('total_score'):=score_home+score_away]
matches= matches[complete.cases(matches)]
matches= matches[score_home>score_away, Result:="Home"]
matches= matches[score_home<score_away, Result:="Away"]
matches= matches[score_home==score_away, Result:="Draw"]

matches[,score_home:=NULL]
matches[,score_away:=NULL]

matches[,ou:=ifelse(total_score>=2.5,"over","under")]


```

Ordering depending on bookmakers

```{r}

setorder(odds,matchId,bookmaker,betType,oddtype,date)
odds2 = odds[,.SD[.N],by=list(matchId,bookmaker,betType,oddtype)]
odds3 = dcast(odds2,matchId~betType+oddtype+bookmaker, value.var = 'odd')

odds4 = odds3[,lapply(.SD,function(x){sum(!is.na(x))})]
```

We have reached the maximum numbers of recurred odds to analyze wisely and selected first 70 ones

```{r}
columnstokeep = names(sort(unlist(odds4),decreasing = T)[1:70])

odds5 = odds3[,columnstokeep,with=F]

odds6 = odds5[complete.cases(odds5)]


odds7=merge(odds6,matches[,c("matchId","ou")],by="matchId")
odds7
pcaodds = prcomp(odds7[,c(-1,-71)],scale. = T)
summary(pcaodds)

```
The model can explain 59% with PC1 and can be increased to 90% by using the first three ones.

#PCA2D

```{r}
autoplot(pcaodds,data=odds7,colour="ou")
```

2D plot of PCA is shown above and colours show if the game is ended over or under.
Regarding the 2D plot, we can not classify bookmakers' odd data.
Let's check on 3D data

#PCA3D

```{r}
pcadt = as.data.table(pcaodds$x)
pcadt

p <- plot_ly(pcadt, x = ~PC1, y = ~PC2, z = ~PC3, color = ~odds7$ou, colors = c('#BF382A', '#0C4B8E')) %>%
  add_markers() %>%
  layout(scene = list(xaxis = list(title = 'PC1'),
                     yaxis = list(title = 'PC2'),
                     zaxis = list(title = 'PC3')))
p


```

Regarding the 3D plot, we can not classify bookmakers' odd data again.

#Manhattan

```{r}
#Distance matrix
manhattan=dist(pcadt,"manhattan")
manhattan[is.na(manhattan)]=0

#Scaling
manhattan=cmdscale(manhattan)
manhattandt=as.data.table(manhattan)
colnames(manhattandt)=c("C1","C2")

m=plot_ly(manhattandt, x = ~C1, y = ~C2, color = ~odds7$ou, colors = c('#BF382A', '#0C4B8E')) %>%
  add_markers() %>%
  layout(scene = list(xaxis = list(title = 'PC1'),
                     yaxis = list(title = 'PC2')))
m
```

#Euclidean

```{r}
#Distance matrix
eu=dist(pcadt)
eu[is.na(eu)]=0

#Scaling
eu=cmdscale(eu)
eudt=as.data.table(eu)
colnames(eudt)=c("C1","C2")

e=plot_ly(eudt, x = ~C1, y = ~C2, color = ~odds7$ou, colors = c('#BF382A', '#0C4B8E')) %>%
  add_markers() %>%
  layout(scene = list(xaxis = list(title = 'PC1'),
                     yaxis = list(title = 'PC2')))
e
```

Results of PCA and MDS are similar. We can not call a difference regarding all the graphs

#Task2
```{r}
require(data.table)
require(TunePareto)
require(glmnet)

testStart=as.Date('2018-11-16')
trainStart=as.Date('2012-07-15')
rem_miss_threshold=0.01 #parameter for removing bookmaker odds with missing ratio greater than this threshold

source('C:\\Users\\a-adsene\\Desktop\\ETM58D\\data_preprocessing.r')
source('C:\\Users\\a-adsene\\Desktop\\ETM58D\\feature_extraction.r')
source('C:\\Users\\a-adsene\\Desktop\\ETM58D\\performance_metrics.r')
source('C:\\Users\\a-adsene\\Desktop\\ETM58D\\train_models.r')


# read data
matches_raw=readRDS("C:\\Users\\a-adsene\\Desktop\\ETM58D\\df9b1196-e3cf-4cc7-9159-f236fe738215_matches.rds")
odd_details_raw=readRDS('C:\\Users\\a-adsene\\Desktop\\ETM58D\\df9b1196-e3cf-4cc7-9159-f236fe738215_odd_details.rds')

# preprocess matches
matches=matches_data_preprocessing(matches_raw)

# preprocess odd data
odd_details=details_data_preprocessing(odd_details_raw,matches)

# extract open and close odd type features from multiple bookmakers
features=extract_features.openclose(matches,odd_details,pMissThreshold=rem_miss_threshold,trainStart,testStart)

# divide data based on the provided dates 
train_features=features[Match_Date>=trainStart & Match_Date<testStart] 
test_features=features[Match_Date>=testStart] 
```
We changed alpha as 0 to eleminate the penalty
```{r}
# run glmnet on train data with tuning lambda parameter based on RPS and return predictions based on lambda with minimum RPS
predictions=train_glmnet(train_features, test_features,not_included_feature_indices=c(1:5), alpha=0,nlambda=50, tune_lambda=TRUE,nofReplications=2,nFolds=10,trace=T)
```
Here are the results;
```{r}
predictions
```